\documentclass
[
    twoside,                 % The thesis is formatted like a book. That is, odd and even pages are handled differently.
    openright,               % Starts a new chapter on an odd page number (right side).
    cleardoublepage = empty, % Clear pages inserted in order to have new chapters appear on odd pages are formatted with an empty style.
    fontsize = 11 pt,        % The size of the font.
    french,                % Support for American English.
    captions = tableheading, % Places the correct amount of space when the caption of a table is above the table.
    numbers = noenddot,      % Does not use a period at the end of numbered titles, such as sections or figures.
    footheight = 35 pt,      % Defines the height of the foot. Due to the line, it needs extra height.
%    draft,                   % Only displays boxes of figures. This option is useful if compilation takes a long time.
]
{scrbook}


\input{settings/document} % Contains commands that are used for certain information that is printed.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Please adjust your options here. %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    % This section contains commands with important data for your thesis. Please adjust them in order for the document to be printed correctly.

    % Defines the length of the amount that a printed page will be cut from EACH side (including the inner side). This option only takes effect with \printVersiontrue and \professionalPrinttrue.
    \extraBorder{3 mm}

    % Shifts the inner margin outward by the amount specified. When the book is bound, part of the page will not be seen anymore. This option compensates for this loss. It only takes effect with \printVersiontrue.
    \bindingCorrection{6 mm}

    % Use the following command if this is a master thesis.

%    \printVersiontrue      % Use this value if you want to prepare your thesis for physical printing. In this case, links will not be colored. Without \professionalPrinttrue, the content will be moved outward by the binding correction, increasing the inner margin and decreasing the outer margin.
%    \professionalPrinttrue % Use this value if you want to have extra border for cutting and are not bound to paper formats (like A4). This option will increase the page size by the extra border on every side plus the binding correction once for the width. It only takes effect in combination with \printVersiontrue.
%    \fancyTheoremsfalse  % Use this value if you want to use the classical theorem style, where the text is italic. Further, with this style, the QED symbol is colorless.
%    \boldNumberSetsfalse % Use this value if you want variables for number sets (like N or R) to appear in blackboard bold rather than bold.

    % The title of the thesis.
    \myTitle{Conception, implémentation et preuve d'un service de transfert de flot d'exécution au sein d'un noyau de système d'exploitation}

    % The author’s name.
    \myName{Florian Vanhems}

    % The author’s program.
    \myProgram{Informatique}

    % A short summary of the thesis. These information will be used for the PDF meta data.
    \mySubject{
The work described in this document is related to formal proofs on operating systems.
The first breakthrough in the domain was the SeL4 project ; demonstrating that producing a complete proof on a microkernel was achievable, albeit very costly. In order to bring the proof's cost down, the CertikOS project showcased a more layered and modular approach, leveraging refinements.
The Pip kernel team tackled the problem from the opposite side, focusing on minimalism, using a shallow embedding methodology and getting rid of refinement altogether. This thesis' contributions are more specifically tied to the Pip kernel.
Previous work on the Pip protokernel focused on providing an isolation proof to Pip's services manipulating the system's memory. Yet, another critical aspect of the kernel -- handling the execution flow transfer from a partition to another -- remained to be designed.
The first contribution of this thesis outlines the design of a single service able to handle all possible control flow transfers in a system ; namely interrupts, faults and explicit control flow transfers. The design focuses on minimalism and code factorization in order to reduce the overall proof effort. An implementation of the service is provided for the Pip kernel. We believe the idea behind the service is general enough to be implemented in other kernels and other architectures.
The second contribution outlined in this thesis is the first formally proven correct userland implementation of an Earliest Deadline First scheduler for arbitrary jobs. The formal proof guarantees that the scheduler's election function respects the Earliest Deadline First scheduling policy, and is guaranteed to be optimal on mono-processor systems. This proof was partly conducted using Pip's usual methodology, leveraging a shallow embedding of the scheduler's code in Coq and a state monad. Nonetheless, while the Pip kernel properties were proven directly, the presented scheduler proofs include three refinement levels ; from the scheduling policy to the actual implementation. Furthermore, the scheduler uses the previously described service in order to pass the control flow to partitions and regain the execution flow through interrupts, showcasing its usability and versatility.
The last contribution presented in this thesis is a proof of concept severing Pip's isolation model from its code. This isolation model independance allows to build alternative models designed to reason on new properties while limiting the proof effort. As such, this contribution opens new research perspectives that were previously too costly to consider. Nonetheless, this proof of concept does not bring the same level of confidence on the composition of properties about the code formally proven on different models.
}

    % Some keywords of the thesis. These information will be used for the PDF meta data. Please use | as a separator and try to avoid commas.
    \myKeywords{Ph.D. thesis | manuscrit de thèse | computer science | informatique | context switch | commutation de contexte | scheduler | ordonnanceur | scheduling | ordonnancement | formal proof | preuve formelle | Coq | proof assistant | assistant de preuve | Earliest Deadline First | models | modèles | operating system | système d'exploitation | kernel | noyau | verification | vérification}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% End of options to adjust. %%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\input{settings/thesis-format}                        % Contains commands that define the general format and layout of the thesis.
\input{settings/bibliography-format}       % Contains commands for the layout of the bibliography.
\input{settings/packages}     % Contains the packages that this template provides.
\input{settings/commands}      % Contains newly defined commands useful for mathematics.

% This is the thesis. The front matter as well as the references should not be changed. The main matter can be edited freely.
\begin{document}


    \frontmatter
    \input{content/front-page}

    \pagestyle{plain}

    \addchap{Abstract}
    \input{content/abstract}

    % Temporary switch of language for the abstract

    \selectlanguage{english}
    \addchap{Abstract}
    \input{content/en-abstract}

    % Switch back to French for the rest of the document

    \selectlanguage{french}
    % Unfortunately we have to redefine the parts commands because the language switch has redefined our redefined commands    
    \input{settings/fr-part-fix}

    \addchap{Remerciements}
    \input{content/remerciements}

    \setuptoc{toc}{totoc}
    \tableofcontents

    \pagestyle{headings}
    \mainmatter

    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %% Please add the content of your thesis here. %%
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \part{Corps du document}
    
    \chapter{Introduction}
    \input{content/introduction}

    \input{content/state-of-the-art}

    \input{content/service}

    \input{content/scheduler}

    \input{content/additional-property}

    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %% End of adding your content. %%
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


    % Add the following chapters not to the current ›part‹ but one level above instead.
    \makeatletter
        \def\toclevel@chapter{-1}
        \def\toclevel@section{0}
    \makeatother

    \chapter{Conclusion}
    \input{content/conclusion}

    % Following are the files and commands for the bibliography and the author’s publications.
    \pagestyle{plain}

    \renewcommand*{\bibfont}{\small}
    \printbibheading
    \addcontentsline{toc}{chapter}{Bibliographie}
    \printbibliography[heading = none]

    \appendix

    \input{content/appendix.tex}
\end{document}
